---
title: "FM Data Wrangling"
author: "Sara"
date: "May 10, 2018"
output:
  html_document:
    df_print: paged
---
### The following file describes the main steps involved in wrangling the Farmer's Market data set from the United States Department of Agriculture (USDA). The file can be found at the link https://www.ers.usda.gov/data-products/food-environment-atlas/data-access-and-documentation-downloads/.

### The raw data was imported and named.
```{r fmdata1, echo=TRUE}
fmdata <- read.csv("fe_local_original.csv")
```

### Data that do not add value to the analysis were removed. This included columns not relevant to farmer's markets, total number of farmer's markets in each county or county equivalent, percentages of each kind of market, and percent change from 2009 to 2016.

```{r fmdata2, echo=TRUE}
fmdata <- fmdata[, -c(4:19,22,24,26,28,30,32,34,36,38,40:100)]
```

### Titles of the remaining columns were changed to make them easier to understand.

#### Data Definitions

* FMRKTPTH09 <- Farmer's markets per 1,000 people in each county in 2009
* FMRKTPTH16 <- Farmer's markets per 1,000 people in each county in 2016
* FMRKT_SNAP16 <- Number of farmer's markets in each county that accepted Supplemental Nutrition Assistance Program (SNAP) benefits in 2016.
* FMRKT_WIC16 <- Number of farmer's markets in each county that accepted Women, Infants, and Children (WIC) benefits in 2016.
* FMRKT_WICCASH16 <- Number of farmer's markets in each county that accepted WIC Cash Value Vouchers (WICCash) in 2016.
* FMRKT_SFMNP16 <- Number of farmer's markets in each county that accepted Senior Farmers' Market Nutrition Program (SFMNP) benefits in 2016.
* FMRKT_CREDIT16 <- Number of farmer's markets in each county that accepted credit cards in 2016.
* FMRKT_FRVEG16 <- Number of farmer's markets in each county that sold fruits and vegetables in 2016.
* FMRKT_ANMLPROD16 <- Number of farmer's markets in each county that sold meat and dairy in 2016.
* FMRKT_BAKED16 <- Number of farmer's markets in each county that sold baked goods in 2016.
* FMRKT_OTHERFOOD16 <- Number of farmer's markets in each county that sold other food products in 2016.

```{r fmdata3, echo=TRUE}
colnames(fmdata)[colnames(fmdata)=="FMRKTPTH09"] <- "Markets2009"
colnames(fmdata)[colnames(fmdata)=="FMRKTPTH16"] <- "Markets2016"
colnames(fmdata)[colnames(fmdata)=="FMRKT_SNAP16"] <- "SNAP"
colnames(fmdata)[colnames(fmdata)=="FMRKT_WIC16"] <- "WIC"
colnames(fmdata)[colnames(fmdata)=="FMRKT_WICCASH16"] <- "WICCash"
colnames(fmdata)[colnames(fmdata)=="FMRKT_SFMNP16"] <- "Senior"
colnames(fmdata)[colnames(fmdata)=="FMRKT_CREDIT16"] <- "CreditCards"
colnames(fmdata)[colnames(fmdata)=="FMRKT_FRVEG16"] <- "Produce"
colnames(fmdata)[colnames(fmdata)=="FMRKT_ANMLPROD16"] <- "Animal"
colnames(fmdata)[colnames(fmdata)=="FMRKT_BAKED16"] <- "Baked"
colnames(fmdata)[colnames(fmdata)=="FMRKT_OTHERFOOD16"] <- "OtherFood"
```

### Numbers in Markets2009 and Markets2016 were converted to percentages rounded to the nearest thousandth for readability and distinguishability.

```{r fmdata4, echo=TRUE}
fmdata$Markets2009 <- round(fmdata$Markets2009, 3)
fmdata$Markets2016 <- round(fmdata$Markets2016, 3)
```

### Columns showing the different kinds of markets per 1,000 people in 2016 were added.

```{r fmdata5, echo=TRUE}
fmdata$SNAP1000 <- fmdata$SNAP*fmdata$Markets2016
fmdata$WIC1000 <- fmdata$WIC*fmdata$Markets2016
fmdata$WICCash1000 <- fmdata$WICCash*fmdata$Markets2016
fmdata$Senior1000 <- fmdata$Senior*fmdata$Markets2016
fmdata$CreditCards1000 <- fmdata$CreditCards*fmdata$Markets2016
fmdata$Produce1000 <- fmdata$Produce*fmdata$Markets2016
fmdata$Animal1000 <- fmdata$Animal*fmdata$Markets2016
fmdata$Baked1000 <- fmdata$Baked*fmdata$Markets2016
fmdata$OtherFood1000 <- fmdata$OtherFood*fmdata$Markets2016
```

### A column showing the difference between Markets2009 and Markets2016 was added.

```{r fmdata6, echo=TRUE}
fmdata$Difference <- fmdata$Markets2016 - fmdata$Markets2009
```

### A column showing whether or not farmer's markets increased their presence in counties between 2009 and 2016 was added.

```{r fmdata7, echo=TRUE}
fmdata$Increased <- ifelse(fmdata$Difference > '0',1,0)
```

### Columns with dummy variables for the different kinds of farmer's markets were added.

```{r fmdata8, echo=TRUE}
fmdata$Present_2009 <- ifelse(fmdata$Markets2009== '0',0,1)
fmdata$Present_2016 <- ifelse(fmdata$Markets2016== '0',0,1)
fmdata$Accepts_SNAP <- ifelse(fmdata$SNAP== '0',0,1)
fmdata$Accepts_WIC <- ifelse(fmdata$WIC== '0',0,1)
fmdata$Accepts_WICCash <- ifelse(fmdata$WICCash== '0',0,1)
fmdata$Accepts_Senior <- ifelse(fmdata$Senior== '0',0,1)
fmdata$Accepts_CreditCards <- ifelse(fmdata$CreditCards== '0',0,1)
fmdata$Sells_Produce <- ifelse(fmdata$Produce== '0',0,1)
fmdata$Sells_Animal <- ifelse(fmdata$Animal== '0',0,1)
fmdata$Sells_Baked <- ifelse(fmdata$Baked== '0',0,1)
fmdata$Sells_OtherFood <- ifelse(fmdata$OtherFood== '0',0,1)
```

### Columns with dummy variables for the states were added.

```{r fmdata9, echo=TRUE}
fmdata$Alabama <- ifelse(fmdata$State== 'AL',1,0)
fmdata$Alaska <- ifelse(fmdata$State== 'AK',1,0)
fmdata$Arizona <- ifelse(fmdata$State== 'AZ',1,0)
fmdata$Arkansas <- ifelse(fmdata$State== 'AR',1,0)
fmdata$California <- ifelse(fmdata$State== 'CA',1,0)
fmdata$Colorado <- ifelse(fmdata$State== 'CO',1,0)
fmdata$Connecticut <- ifelse(fmdata$State== 'CT',1,0)
fmdata$Delaware <- ifelse(fmdata$State== 'DE',1,0)
fmdata$DistrictofColumbia <- ifelse(fmdata$State== 'DC',1,0)
fmdata$Florida <- ifelse(fmdata$State== 'FL',1,0)
fmdata$Georgia <- ifelse(fmdata$State== 'GA',1,0)
fmdata$Hawaii <- ifelse(fmdata$State== 'HI',1,0)
fmdata$Idaho <- ifelse(fmdata$State== 'ID',1,0)
fmdata$Illinois <- ifelse(fmdata$State== 'IL',1,0)
fmdata$Indiana <- ifelse(fmdata$State== 'IN',1,0)
fmdata$Iowa <- ifelse(fmdata$State== 'IA',1,0)
fmdata$Kansas <- ifelse(fmdata$State== 'KS',1,0)
fmdata$Kentucky <- ifelse(fmdata$State== 'KY',1,0)
fmdata$Louisiana <- ifelse(fmdata$State== 'LA',1,0)
fmdata$Maine <- ifelse(fmdata$State== 'ME',1,0)
fmdata$Maryland <- ifelse(fmdata$State== 'MD',1,0)
fmdata$Massachusetts <- ifelse(fmdata$State== 'MA',1,0)
fmdata$Michigan <- ifelse(fmdata$State== 'MI',1,0)
fmdata$Minnesota <- ifelse(fmdata$State== 'MN',1,0)
fmdata$Mississippi <- ifelse(fmdata$State== 'MS',1,0)
fmdata$Missouri <- ifelse(fmdata$State== 'MO',1,0)
fmdata$Montana <- ifelse(fmdata$State== 'MT',1,0)
fmdata$Nebraska <- ifelse(fmdata$State== 'NE',1,0)
fmdata$Nevada <- ifelse(fmdata$State== 'NV',1,0)
fmdata$NewHampshire <- ifelse(fmdata$State== 'NH',1,0)
fmdata$NewJersey <- ifelse(fmdata$State== 'NJ',1,0)
fmdata$NewMexico <- ifelse(fmdata$State== 'NM',1,0)
fmdata$NewYork <- ifelse(fmdata$State== 'NY',1,0)
fmdata$NorthCarolina <- ifelse(fmdata$State== 'NC',1,0)
fmdata$NorthDakota <- ifelse(fmdata$State== 'ND',1,0)
fmdata$Ohio <- ifelse(fmdata$State== 'OH',1,0)
fmdata$Oklahoma <- ifelse(fmdata$State== 'OK',1,0)
fmdata$Oregon <- ifelse(fmdata$State== 'OR',1,0)
fmdata$Pennsylvania <- ifelse(fmdata$State== 'PA',1,0)
fmdata$RhodeIsland <- ifelse(fmdata$State== 'RI',1,0)
fmdata$SouthCarolina <- ifelse(fmdata$State== 'SC',1,0)
fmdata$SouthDakota <- ifelse(fmdata$State== 'SD',1,0)
fmdata$Tennessee <- ifelse(fmdata$State== 'TN',1,0)
fmdata$Texas <- ifelse(fmdata$State== 'TX',1,0)
fmdata$Utah <- ifelse(fmdata$State== 'UT',1,0)
fmdata$Vermont <- ifelse(fmdata$State== 'VT',1,0)
fmdata$Virginia <- ifelse(fmdata$State== 'VA',1,0)
fmdata$Washington <- ifelse(fmdata$State== 'WA',1,0)
fmdata$WestVirginia <- ifelse(fmdata$State== 'WV',1,0)
fmdata$Wisconsin <- ifelse(fmdata$State== 'WI',1,0)
fmdata$Wyoming <- ifelse(fmdata$State== 'WY',1,0)
```

### Columns with dummy variables for the regions were added. Examining these columns showed the values as characters, so they were converted to numeric.

```{r fmdata10, echo=TRUE}
fmdata$Midwest <- paste(fmdata$Illinois+fmdata$Indiana+fmdata$Iowa+fmdata$Kansas+fmdata$Michigan+fmdata$Minnesota+fmdata$Missouri+fmdata$Nebraska+fmdata$NorthDakota+fmdata$Ohio+fmdata$SouthDakota+fmdata$Wisconsin)
fmdata$Northeast <- paste(fmdata$Connecticut+fmdata$Delaware+fmdata$DistrictofColumbia+fmdata$Maine+fmdata$Maryland+fmdata$Massachusetts+fmdata$NewHampshire+fmdata$NewJersey+fmdata$NewYork+fmdata$Pennsylvania+fmdata$RhodeIsland+fmdata$Vermont)
fmdata$South <- paste(fmdata$Alabama+fmdata$Arkansas+fmdata$Florida+fmdata$Georgia+fmdata$Kentucky+fmdata$Louisiana+fmdata$Mississippi+fmdata$NorthCarolina+fmdata$Oklahoma+fmdata$SouthCarolina+fmdata$Tennessee+fmdata$Texas+fmdata$Virginia+fmdata$WestVirginia)
fmdata$West <- paste(fmdata$Alaska+fmdata$Arizona+fmdata$California+fmdata$Colorado+fmdata$Hawaii+fmdata$Idaho+fmdata$Montana+fmdata$Nevada+fmdata$NewMexico+fmdata$Oregon+fmdata$Utah+fmdata$Washington+fmdata$Wyoming)

fmdata$Midwest <- as.numeric(as.character(fmdata$Midwest))
fmdata$Northeast <- as.numeric(as.character(fmdata$Northeast))
fmdata$South <- as.numeric(as.character(fmdata$South))
fmdata$West <- as.numeric(as.character(fmdata$West))
```

### Cells with "NA" values are in the same rows as Markets2016 values with zeroes, so they were replaced with zeroes to make all values numeric.

```{r fmdata11, echo=TRUE}
fmdata[is.na(fmdata)] <- 0
```

### A logarithmic transformation function was added to normalize the data consisting of markets per 1,000 people.

```{r fmdata12, echo=TRUE}
signedlog10 = function(x) {
  ifelse(abs(x) <= 1, 0, sign(x)*log10(abs(x)))
}
```

### The logarithmic transformation function was applied to the data of markets per 1,000 people and then the data was normalized and rounded to the nearest thousandth.

```{r fmdata13, echo=TRUE}
fmdata$Markets_2009_norm <- signedlog10(fmdata$Markets2009)
fmdata$Markets_2009_norm <- rnorm(fmdata$Markets_2009_norm)
fmdata$Markets_2009_norm <- round(fmdata$Markets_2009_norm, 3)

fmdata$Markets_2016_norm <- signedlog10(fmdata$Markets2016)
fmdata$Markets_2016_norm <- rnorm(fmdata$Markets_2016_norm)
fmdata$Markets_2016_norm <- round(fmdata$Markets_2016_norm, 3)

fmdata$SNAP_norm <- signedlog10(fmdata$SNAP1000)
fmdata$SNAP_norm <- rnorm(fmdata$SNAP_norm)
fmdata$SNAP_norm <- round(fmdata$SNAP_norm, 3)

fmdata$WIC_norm <- signedlog10(fmdata$WIC1000)
fmdata$WIC_norm <- rnorm(fmdata$WIC_norm)
fmdata$WIC_norm <- round(fmdata$WIC_norm, 3)

fmdata$WICCash_norm <- signedlog10(fmdata$WICCash1000)
fmdata$WICCash_norm <- rnorm(fmdata$WICCash_norm)
fmdata$WICCash_norm <- round(fmdata$WICCash_norm, 3)

fmdata$Senior_norm <- signedlog10(fmdata$Senior1000)
fmdata$Senior_norm <- rnorm(fmdata$Senior_norm)
fmdata$Senior_norm <- round(fmdata$Senior_norm, 3)

fmdata$CreditCards_norm <- signedlog10(fmdata$CreditCards1000)
fmdata$CreditCards_norm <- rnorm(fmdata$CreditCards_norm)
fmdata$CreditCards_norm <- round(fmdata$CreditCards_norm, 3)

fmdata$Produce_norm <- signedlog10(fmdata$Produce1000)
fmdata$Produce_norm <- rnorm(fmdata$Produce_norm)
fmdata$Produce_norm <- round(fmdata$Produce_norm, 3)

fmdata$Animal_norm <- signedlog10(fmdata$Animal1000)
fmdata$Animal_norm <- rnorm(fmdata$Animal_norm)
fmdata$Animal_norm <- round(fmdata$Animal_norm, 3)

fmdata$Baked_norm <- signedlog10(fmdata$Baked1000)
fmdata$Baked_norm <- rnorm(fmdata$Baked_norm)
fmdata$Baked_norm <- round(fmdata$Baked_norm, 3)

fmdata$OtherFood_norm <- signedlog10(fmdata$OtherFood1000)
fmdata$OtherFood_norm <- rnorm(fmdata$OtherFood_norm)
fmdata$OtherFood_norm <- round(fmdata$OtherFood_norm, 3)
```

### Export data to new cleaned CSV file.

```{r fmdata14, echo=TRUE}
write.csv(fmdata, "fe_local_clean.csv")
```